A quantum parallel Markov chain Monte Carlo

### 2023-01-17 main.tex
black hole caption

Once one has effectively parallelized the target evaluations $\pi(\ttheta_{p})$, the new computational bottleneck is the calculations required to obtain the $P+1$ proposal density terms $q(\ttheta_{p},\Ttheta_{-p})$, for $p=0,\dots,P$.  Under an independent proposal mechanism with a proposal distribution of the form $q(\ttheta_0,\Ttheta_{-0})=\prod_{p=1}^Pq(\ttheta_0,\ttheta_{p})$, the acceptance probabilities \eqref{eq:probs} require the $\order{P^2}$ evaluation of the $P+1\choose 2$ terms $q(\ttheta_{p},\ttheta_{p'})$.  To avoid such calculations, we follow \citet{holbrook2021generating} and use symmetric proposal distributions for which $q(\ttheta_{p},\Ttheta_{-p}) = q(\ttheta_{p'},\Ttheta_{-p'})$ for $p,p'=0,\dots,P$ (Section \ref{sec:sym}).

@@

\begin{figure}[!t]
	\centering
	\scalebox{0.9}{
		\begin{minipage}{1\textwidth}
			\begin{algorithm}[H]
	\caption{Quantum parallel MCMC}\label{alg:qpMCMC}
	\KwData{Initial Markov chain state $\ttheta^{(0)}$; total length of Markov chain $S$; total number of proposals per iteration $P$; routine for evaluating target density $\pi(\cdot)$; routines for drawing random samples from a $D$-dimensional multivariate Gaussian $Normal_D(\boldsymbol{\mu},\SSigma)$ and the standard Gumbel distribution $Gumbel(0,1)$.}
	\KwResult{A Markov chain $\ttheta^{(1)}, \dots, \ttheta^{(S)}$.}
	\For{$s \in\{1,\dots,S\}$}{
		$\ttheta_0 \gets \ttheta^{(s-1)}$\;
		$z_{0} \gets Gumbel(0,1)$\;
		$\bar{\ttheta} \gets Normal_D(\ttheta_0,\SSigma)$; \hspace{14em} \Comment{ \hspace{0.3em}     Proposal \eqref{eq:p1}}      
		\For{$p \in \{1,\dots,P\}$}{
			$\ttheta_p \gets Normal_D(\bar{\ttheta},\SSigma)$\;
			$z_{p} \gets Gumbel(0,1)$\;
		}
		$\hat{p} \gets \argmin_{p=0,\dots,P}\Big(f(p):= -\big(z_{p}+ \log\pi(\ttheta_{p}) \big)\Big)$\;  \hspace{10em}\Comment{\hspace{0.5em} $\order{\sqrt{P}}$ Alg \ref{alg:min} with warm-start at $p=0$.}
		$\ttheta^{(s)} \gets \ttheta_{\hat{p}}$\;
	}
	\Return{$\ttheta^{(1)}, \dots, \ttheta^{(S)}$}\ .

@@

Algorithm \ref{alg:qpMCMC} presents the details of QPMCMC.  The algorithm uses conventional (perhaps parallel) computing almost entirely, excluding the one line that calls the quantum minimization algorithm presented in Algorithm \ref{alg:min}. 
This sub-routine seeks to find the minimizer of the function
\begin{align*}
f(p)= -\big(z_{p}+ \log\pi(\ttheta_{p}) \big) \quad \mbox{over} \quad p=0,1,\dots, P,
\end{align*} 
which combines the numerators of \eqref{eq:probs}, the Gumbel-max trick and a trivial negation.  As discussed in Section \ref{sec:qmin}, the sub-routine requires only $\mathcal{O}(\sqrt{P})$ oracle evaluations and, therefore, only $\mathcal{O}(\sqrt{P})$ evaluations of the target density $\pi(\cdot)$ using a quantum circuit.  In theory, a quantum computer can perform the same computations as a classical computer, but the efficiency of the target evaluations would depend on a number of factors related to the structure of the density itself, the availability of fast quantum random access memory \citep{giovannetti2008quantum} and the ability of large numbers of quantum gates to act in concert with negligible noise \citep{kielpinski2002architecture,erhard2019characterizing}.

@@

\subsubsection{Uniform proposals}\label{sec:unif}

@@

 \begin{figure}[!t]
	\centering
	\includegraphics[width=0.7\linewidth]{Ising2dFig.pdf}
	\caption{Convergence of log-posterior for a parallel MCMC sampler targeting an Ising model with uniform correlation $\lambda=1$ and no external field. All chains begin at minimal probability `checkerboard' state. Larger proposal counts allow discovery of higher probability configurations, albeit with diminishing returns.}\label{fig:ising2d}
\end{figure}

@@

 \begin{figure}[!t]
	\centering
	\includegraphics[width=0.9\linewidth]{blackHole.png}
	\caption{The }\label{fig:blackHole}
\end{figure}

\subsection{Note on simulations}

@@

>>>>

Once one has effectively parallelized the target evaluations $\pi(\ttheta_{p})$, the new computational bottleneck is the calculations required to obtain the $P+1$ proposal density terms $q(\ttheta_{p},\Ttheta_{-p})$, for $p=0,\dots,P$.  Under an independent proposal mechanism with a proposal distribution of the form $q(\ttheta_0,\Ttheta_{-0})=\prod_{p=1}^Pq(\ttheta_0,\ttheta_{p})$, the acceptance probabilities \eqref{eq:probs} require the $\order{P^2}$ evaluation of the $P+1\choose 2$ terms $q(\ttheta_{p},\ttheta_{p'})$.  To avoid such calculations, we follow \citet{holbrook2021generating} and use symmetric proposal distributions for which $q(\ttheta_{p},\Ttheta_{-p}) = q(\ttheta_{p'},\Ttheta_{-p'})$ for $p,p'=0,\dots,P$.

@@

\begin{figure}[!t]
	\centering
	\scalebox{0.9}{
		\begin{minipage}{1\textwidth}
			\begin{algorithm}[H]
	\caption{Quantum parallel MCMC for a continuously-valued target}\label{alg:qpMCMC}
	\KwData{Initial Markov chain state $\ttheta^{(0)}$; total length of Markov chain $S$; total number of proposals per iteration $P$; routine for evaluating target density $\pi(\cdot)$; routines for drawing random samples from a $D$-dimensional multivariate Gaussian $Normal_D(\boldsymbol{\mu},\SSigma)$ and the standard Gumbel distribution $Gumbel(0,1)$.}
	\KwResult{A Markov chain $\ttheta^{(1)}, \dots, \ttheta^{(S)}$.}
	\For{$s \in\{1,\dots,S\}$}{
		$\ttheta_0 \gets \ttheta^{(s-1)}$\;
		$z_{0} \gets Gumbel(0,1)$\;
		$\bar{\ttheta} \gets Normal_D(\ttheta_0,\SSigma)$; \hspace{14em} \Comment{ \hspace{0.3em}     Proposal \eqref{eq:p1}}      
		\For{$p \in \{1,\dots,P\}$}{
			$\ttheta_p \gets Normal_D(\bar{\ttheta},\SSigma)$\;
			$z_{p} \gets Gumbel(0,1)$\;
		}
		$\hat{p} \gets \argmin_{p=0,\dots,P}\Big(f(p):= -\big(z_{p}+ \log\pi(\ttheta_{p}) \big)\Big)$\;  \hspace{10em}\Comment{\hspace{0.5em} $\order{\sqrt{P}}$ Alg \ref{alg:min} with warm-start at $p=0$.}
		$\ttheta^{(s)} \gets \ttheta_{\hat{p}}$\;
	}
	\Return{$\ttheta^{(1)}, \dots, \ttheta^{(S)}$}\ .

@@

Algorithm \ref{alg:qpMCMC} presents the details of QPMCMC for a continuously-valued target.  The algorithm uses conventional (perhaps parallel) computing almost entirely, excluding the one line that calls the quantum minimization algorithm presented in Algorithm \ref{alg:min}. 
This sub-routine seeks to find the minimizer of the function
\begin{align}\label{eq:objective}
f(p)= -\big(z_{p}+ \log\pi(\ttheta_{p}) \big) \quad \mbox{over} \quad p=0,1,\dots, P,
\end{align} 
which combines the numerators of \eqref{eq:probs}, the Gumbel-max trick and a trivial negation.  As discussed in Section \ref{sec:qmin}, the sub-routine requires only $\mathcal{O}(\sqrt{P})$ oracle evaluations and, therefore, only $\mathcal{O}(\sqrt{P})$ evaluations of the target density $\pi(\cdot)$ using a quantum circuit.  In theory, a quantum computer can perform the same computations as a classical computer, but the efficiency of the target evaluations would depend on a number of factors related to the structure of the density itself, the availability of fast quantum random access memory \citep{giovannetti2008quantum} and the ability of large numbers of quantum gates to act in concert with negligible noise \citep{kielpinski2002architecture,erhard2019characterizing}.

@@

@@

 \begin{figure}[!t]
	\centering
	\includegraphics[width=0.7\linewidth]{Ising2dFig.pdf}
	\caption{Convergence of log-posterior for a parallel MCMC sampler targeting an Ising model with uniform correlation $\lambda=1$ and no external field. All chains begin at minimal probability `checkerboard' state. Larger proposal counts allow discovery of higher probability configurations.}\label{fig:ising2d}
\end{figure}

@@

 \begin{figure}[!t]
	\centering
	\includegraphics[width=0.9\linewidth]{blackHole.png}
	\caption{On the left is a $4076\times4076$ intensity map of the shadow of supermassive black hole Sagittarius A*  \citep{akiyama2022first}.  On the right is the pixelwise posterior mode of a Bayesian image classification model fit to intensity data.  Within a Metropolis-in-Gibbs routine, quantum parallel MCMC requires less that one-tenth the posterior evaluations required by conventional parallel MCMC.}\label{fig:blackHole}
\end{figure}

@@

\section{Note on simulations}
